### 2015-11-14 paper/paper.tex
fixed cookie issue with volksrant.nl, but it persists with telegraafnl (we may want to skip oveir the site). Spider should now work for the volkskrant.

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\title{Illustrating Text}
\author{Christopher Catton, Partick Murphy, Charles Ung}
\begin{document}
\section{Abstract}
\section{Generating Dataset(s)}
In our study we created our own dataset(s) in order to generate a dataset containing the data we need. We generated the dataset by crawling and scraping separate news sites. Our data is tuple of the source, date, and text of an article.
Instead of gathering data related to images, we use google image search to generate a set of images related to a set and subsets of the words we extract from a text. In other related studies this task can be done by adding images that appear with a text, but we chose to use google image search to reduce the amount of scraping we need to do to generate our dataset.
\end{document}