### 2015-12-10 paper/paper.tex
minor changes to textAnalysis and much more progress on the paper. Biggest thing to do is the related work and bringing in some images from the program.

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{comment}
\title{Illustrating Text}
\author{Christopher Catton, Partick Murphy, Charles Ung}
\date{December 10, 2015}
\begin{document}
\maketitle
\section{Introduction}
%discuss the human link -- the relationship of text and visualization -- motivation
\paragraph{}
A text has always been a complicated thing to translate into another language. Language
is not precise and encoding an idea into text is not always perfect. Often, Language 
has a number of nuances brought in by different culture. Someone from one country may
associate luck with red and someone from another country may associate luck with green.
But, there still will be an intersection in the features, the visualization, of what
any two people assign to a word or even a text in general. If it was not the case then
we couldn't translate anything. But, we know that we can translate texts to different
languages and we know that multiple texts can describe the same thing.
\paragraph{}
Often in translation of a text there is some meaning lost. Different cultures view things
differently and different languages encode ideas differently. Everyone tends to visualize 
meaning to some extent though.

%define what we are doing in a basic form
\paragraph{}
In this paper, we take databases of texts retrieved from sources with different languages and
cultures. We use the databases, with no prior understanding of the language or cultures, to
explore whether we can determine a subset of each text and assign it some meaning by obtaining
images for further subsets of the subset generated for the text. In doing so we expect that
we can generate meaning for a text and that we can determine whether two texts encode the
same thing (i.e. \ describe the same event or idea).

%define applications
\paragraph{}
An interesting, but difficult application for this problem is to generate visual translations. 
Providing an image for a small set of texts can aide in translation or even language learning
tasks, but it is dificult to illustrate certain language elements. Take for example elements
such as `a', `but', `or', or even `context'. There are a lot of elements that may not be 
easily transcribed into an image, but others such as `man', `woman', and `building' could
easily be transcribed. There are also more complex elements that can be described such as
the word `luck' or `lottery'. There are difficulties in any approach to translate and 
teaching languages and we believe that generating visual illustrations as we attempt to do 
is a possible application.

\paragraph{}
Generating datasetss for related problems
Another application is generating datasets for other related problems such as trying to 
describe a given image. The task has been tried before and this could potentially generate
a dataset for training a neural network for doing the task. This probably could be best 
achieved by building a dataset for training an object recognition network. In the simplest
form of the problem we are setting out to achieve could result in providing a set of collages
with images keyed to a certain location in each matrix produced from a keyword/phrase set.

\paragraph{}
There are many facets to the given problem and this paper is organize to partition each step. This
paper first discusses the basic method of acheiving results and then details each step. The details
are then followed by our conclusions of our exploration. The last section is a discussion on what 
can be done to improve our results and what modifications can be done to create more complex, 
interesting, and generally better performing models.

\section{Related Work}
Related Work
-- Computer Vision

\section{Basic Method}

\paragraph{}
The basic method or principle for the problem is to take or generate a language model for
each language that is to be used. Then to take a set of texts written in languages that we
have models for and filter them on some level to generate a set or sets of keywords for each
text. We then take a model that converts the keywords into a set of images with a similar
format. Then we perform the very important step of partitioning the images into subsections
between the languages. We then arrive at the last step where we take the set of images and
combine them in a meaningful way to illustrate the texts.

\paragraph{}
In generating the language model for the languages used, we used basic apriori methods to
give us relevant data for words and word pairs. Usage of non-aprori methods in way works
against what we are attempting to do since non-apriori methods may be encoding meaning
of words or word combinations into account, but we are attempting to solve this through
the use of images annotated with keywords and phrases. Non-apriori methods may provide
a better measure of salience then apriori method, but take away from what we are trying
to explore in a finer grained sense. Ideally we should be able to assign meaning to a text
without already knowing details of a language as mentioned priori.

\paragraph{}
The partitioning of images is the task we would say is the most important step. It is where
we believe we can take a keyword or phrase, a word part, from one language and assign it a 
meaning that we can intrinsically understand. Within one language there will be a number of
sets that are more descriptive of the nuances of the meanings of the keywords within the
language and others less so. We determine how much a set represents the nuances of a 
language by seeing if it also exists as a set for another language. The intersection between
the image sets of two texts acts a way of being able, or a measure, to determine whether or 
not two texts are two different encodings of the same thing. If the two texts are encoding 
the same thing, i.e. \ the measure is high, then logically image sets that measure less similar
are still encoding the same material. But, they are also better encodings for their language
since they are also incorporating nuances of the language and culture encoding them.

\paragraph{}
There are more advanced methods for dealing with image sets that should provide much better
results and more flexibility in their use. Those methods are discussed at the end of the paper. 
We felt that usage of simple methods would act as a much better illustration of the problem and
its difficulty. Our decision to use simpler methods does come with a number of difficulties in
itself that may not be difficulties of the advanced methods discussed.

\section{Generating Language Model Dataset(s)}
In our study we created our own dataset(s) in order to generate a dataset containing the data we 
need. We generated the dataset by crawling and scraping separate news sites. These new sites 
include the British Broadcasting Corporation, the New York Times, De Telegraaf, and De Volkskrant. 
Our data is tuple of the source, date, and text of an article. Instead of gathering data related 
to images, we use google image search to generate a set of images related to a set and subsets of 
the words we extract from a text. In other related studies this task can be done by adding images 
that appear with a text, but we chose to use google image search to reduce the amount of scraping 
we need to do to generate our dataset.

\section{Generate Language Models}

\paragraph{}
In our study we create language models for languages of articles that we illustrate. We do this using 
apriori methods and first find frequency for each word and for each set of words with each set being 
of length N or less. We automatically remove punctuation marks and split on each space and parse each
word as part of sets between the size of 1 < N, with N being the max set size. Each set is constructed
with words found by each other. This is the stage we generate frequency overall, and frequency over 
articles for each word and word pair. This stage should be achieved while parsing each article only once.

\paragraph{}
In more detail, we select a number of sites that are written in a specific language and parse the articles
we have already obtained in generating our dataset. Each Article is then processed to obtain the frequency
for each word and phrase of an appropriate size. After each Article is parsed its words and statistics are
added to the language model.

\paragraph{}
We expect the least interesting articles would appear the most frequently across articles in any given 
language. In-frequent word sets over all articles that are frequent in a small set of articles are likely 
to be salient for a given article. It is difficult to establish word and word-set salience even with the 
frequencies. Some words may be infrequent and frequent in articles, but not really descriptive of the 
article in general. In using this stage it is necessary to establish cut-off points for salient and 
non-salient word-sets.

\section{Generating Salient Keyword/Phrase Sets}
Generating salient sets is fairly easy compared with establishing a language model. To get a set of 
salient-words for a given article we provide the statistics for a given article and take the language 
model for the article. We use the language model to remove non-salient words from the article and then 
use some combination of statistics of the word for the article and in the language. It is hard to
determine a good combination of the statistics that represent the salience of a given word or word-set. 
Determining a statistic that represents good salience for a word or word-set is a problem on its own.
\paragraph{}
In this paper we choose a the combination:
$$1-frequencyInArticle(Word)/numberOfWordsInArticle$$
to represent the salience of a word or word-set.
\paragraph{}
We also use two thresholds to remove non-salient words. One is the percent of articles that a word or
word-set appears in and the other is how many occurences does the word appear overall.

\section{Generating Image Sets}
In our project we require a large number of images with similar annotations of keywords for each language 
used. The keywords for certain images may differ for each image, but there should be an intersection among 
the images for a select set of keywords. Since this is difficult to obtain on our own we, in our project, 
turned to google to obtain sets of images for keywords.
\paragraph{}
In generating a number of seperate sets for each set of salient keyword/phrase sets, we hope to be able
to determine whether there are any sets, generated by different texts, that are similar enough to be
considered the same. If they are similar enough it follows that the texts are encoding the same material 
and we can possibly say that if the articles are in different languages then they are approximate 
translations of each other.
\paragraph{}
After downloading images from the Google search image API of salient word sets from the article, the common 
subset of images within these sets must be found. This can be done by finding the intersection of all the 
sets. In our implemnetation, we first created subset that was the intersection of two sets and then 
iteratively performed the intersection of this subset with each other set of images. Since the number of 
images is relatively large, all the images in two sets could not be loaded into memory at once. To deal with 
this issue, the intersection algorithm used was a block-based intersection which loads two blocks of 50 images 
into memory, then compares each image in 1 block to each image in the other block and output images which are 
in both blocks. To compare the images two different comparison tests were used. For two images to be considered 
the same, they must pass either or both of the comparison tests. The first test was a simple pixel by pixel 
comparison, if for each pixel in each image the values of the pixels were equal then the two images are the 
same. The second comparison test was to take the average colour of all the pixels in a different square regions 
of the images and compare them. If the average colours in each of the regions were the same, then the images 
are considered the same.

\section{Generating Illustrations}
The last and final step in our project is how we visualize the image-sets and the keyword salience. Ideally we 
would be able to combine the images into a single images, but that is a difficult task even if we know what 
features and feature-groups a keyword or keyword-set relates to. Instead we opted for creating collage of the 
images within a set. This means that we would generate something similar to what one might expect an output of 
a model for object segmentation. There are no masks of objects, but we might expect that certain positions in 
a matrix to be occupied by a certain object for a salient set.
\paragraph{}
In arranging images in a collage we get around the problem of trying to combine them and possibly provide 
something in way of a story arc. It should show more meaningful visualization of the article at the origin of 
the matrix then in other places, provided that the salience measure is correct. It is also fairly culturally 
ambiguous since a lot of cultures have come to expect items to start in the top-left in recent times. But, 
there are more culturally ambiguous ways to create illustrations as well (which are more difficult).

\section{Conclusion and Problems}
One problem we encountered was that the Google image seach API only allows for 100 searches per day for free, so 
gathering enough images for each of the salient sets was impossible. This in turn has made it impossible to 
get the results we would consider significant. To obtain the results we want we will need access to a database 
that we can obtain a much larger set of images for each salient set. This may be achieved by data-mining for
images based on the salient sets themselves. Though there are possible ways of reducing the requirement of a
large amount of images which are discussed later.
\paragraph{}
Another problem is validation of results. Since we were unable to obtain the results needed, it is not relevant
that we cannot evaluate them. But, it needs to be discussed how we would go about evaluating the results had 
we obtained them. For this task we probably we need to crowd-source in order to have people evaluate the accuracy 
of the results. It isn't out of ordinary to need actual people to be involved the actual evaluation of the results. 
In computer vision a lot of data for evaluating models for generating salience or object segmentation have manually 
generated data from people. For us, we would not need any special tools such as eye-tracking tools to determine 
the accuracy though.

\section{Future Work}
-- better datasets (particularly for images)
-- more complex analysis of images
-- break down of images into features and feature groups

\subsection{Datasets}
Future work needs access to a larger set of images annotated with keywords and phrases. It may be necessary to mine 
for the images and annotations specifically or to crowd source the annotations of mined images. But, it is achievable 
and may provide the results we wished to achieve. One issue we found with google is that the search, provided with two 
words that are nearly identical in meaning, does not provide nearly identical results. Ideally images would have 
similar ranking in a search for words with a similar meaning. We went in expecting there to be a large reuse of images 
between people of different language groups, but our observations now point us in a different direction.

\subsection{Language Model}
A possible extension of how we generate a language model is to take into account the time each article is written and 
published. We expect that certain words are more likely to be published during certain time periods of the year such as 
during holidays (e.g. Christmas, Thanksgiving). Also, articles related to a crisis would likely be published during a 
certain span of time and possibly articles would be published in rememberance of them (e.g. 911).
\paragraph{}
In many languages their is a tendency for certain word-parts to take a certain order which may vary for a number of 
reasons such as whether a phrase is a question or statement. In later iterations of our language model we can take 
this into account and try to determine whether a language has such an order and when or whether it changes during 
certain contexts. We can determine whether there is a word order by determining if two or more words appear 
along-side each other relatively frequently.
\paragraph{}
Another concern about word-parts is that certain parts, such as nouns, may be considered more salient for illustration. 
It may be possible to determine whether a particular word is a noun, verb, etc. \ by considering their frequency 
along-side closely occuring words and their order. But, this is a difficult task on its own.
\paragraph{}
Another consideration is that some languages compound words together or have prefixes and suffixes attached to words in certain contexts. 
In order for us to determine whether a word occurs within a compound word would be to check whether it appears alone, 
it's frequency overall (the word `a' is likely to appear in a lot of words, but it is not a subword), and its frequency as a subword.
\paragraph{}
The task of getting relevant salient sets of words isn't necessarily resrticed to apriori methods or even traditional 
data-mining techniques. In a way using non-apriori methods may detract from ascertaining the meaning from the image sets 
we find. But, their results could still be considered interesting and relevant. It would be an interesting task to design 
a neural network to determine word salience or even to get it to create the illustrations. Some machine learning and a 
training set to determine thresholds for what is and is not salient could go a long ways. The thresholds would probably 
maintain relevance across different methods of solving the problem, but right now we are only guessing and manually 
adjusting them.

\subsection{Image Analysis and Partitioning of Image Sets}
A possible extension of this stage would be to incorporate machine learning and computer vision to be able to find 
features and feature-groups within an image. Using features and feature-groups we may be able to find which features 
or feature-groups relate to each keyword or keyword-set and pick images based on this. This step is already improved 
if we are able to match multiple modifications of a single image and throwing away the modified versions. Looking at 
an image on a feature basis allows us to generate a better set of images that represent the keywords better.

\subsection{Illustration}
-- Composite Images based on keywords
-- Exploration of culturally ambiguous collages (Left-to-Right Right-to-Left)
-- Image Modification based Salience and Computer Vision

\end{document}